Online Social Cohesion Reflects Real-World Group Action in Syria during the Arab Spring

### 2020-12-23 manuscript.Rmd
Update intro pt. 2

In estimating cohesion in Twitter posts, tweets that share content words
yield higher cohesion values, even among speakers with opposing
views on the subject matter. Repetition of terms and concepts is captured by
reiteration, which is a type of lexical cohesion in which a subsequent item
repeats a previous item, or refers to it through synonymous language
[@halliday1976cohesion]. In social media, we can conceive of retweets (i.e., reposts
or forwards of previous messages) as a form of reiteration
that increase message cohesion. Since retweets replicate verbatim some part of
another tweet, these tweets would also have higher cohesion. Hashtags function
similarly and are included in our analysis.

@@

@@

Fed by lexical cohesion, interlocutor cohesion ebbs and flows, contextualized by
changing real world events. Interlocutor cohesion may increase preceding an
in-person social mobilization event, alerting a repressive government that
collective mobilization is happening. Interlocutor cohesion may also respond to
exogenous events such as spontaneous, unprovoked militarized conflict.

@@

We expect to see increased linguistic cohesion among people participating in
online discussions about an idea, event, individual, or movement. Given that
Syrian protesters coordinated their activity through Twitter by discussing their
own events (e.g., planned protests, active demonstrations) and their reactions
to events happening around them (e.g., government reactions, international
action or inaction), the cohesion in social media should correlate with the
social cohesion of the group [@zeitzoff2011using].

@@

## The Present Work
Social media has become a primary platform for social movements and activism
across the world, but the extent to which virtual activism reflects and drives
real-world mobilization is still largely unknown. In the current work, we
conceptualize the physical and virtual movements as a dynamical system, both
constantly interacting and influencing the other. We aim to study this by
analyzing the real-world events and online social cohesion during nearly three
months of the Arab Spring. Both real- and virtual-world collective coordination
were powerful tools in the events of the Arab Spring and are undeniably linked.
This study aims to explore the dynamics of Twitter users in Syria, who used the
social media platform to raise awareness about the Arab Spring and bring about
mass mobilization.

@@

This work is a reproducible manuscript. All code and analyses can be found at:
www.github.com/mchiovaro/arab-spring.

>>>>

In estimating cohesion in Twitter posts, tweets that share content words
yield higher cohesion values, even among speakers with opposing
views on the subject matter. Repetition of terms and concepts is captured by
reiteration, which is a type of lexical cohesion in which a subsequent item
repeats a previous item, or refers to it through synonymous language
[@halliday1976cohesion]. We include retweets (i.e., reposts or forwards of previous messages) because they are a form of reiteration that increases message cohesion. They indicate that individuals are “singing to the same tune” by amplifying the original message. Since retweets replicate verbatim another tweet (sometimes with the addition of commentary or supplementary messages), these tweets would also have higher cohesion.

@@

We also include hashtags in our analyses. Hashtags are a way of identifying similar topics (e.g., people, events, places, themes) across users and across tweets by typing a string of characters preceded by a “#” symbol. Users may choose to include as many hashtags (up to the total character limit of the tweet) as they would like, and they may choose to create their own or adopt hashtags from other users or groups. The use of the same hashtags would increase cohesion; however, because hashtags are often included or embedded within longer messages, simply using the same hashtags would not be a guarantee of higher cohesion if the rest of the words in the tweets do not overlap.

@@

Fed by lexical cohesion, interlocutor cohesion ebbs and flows, contextualized by
changing real world events. Interlocutor cohesion may increase preceding an
in-person social mobilization event, alerting a repressive government that
collective mobilization is happening. It may also respond to
exogenous events such as spontaneous, unprovoked militarized conflict.

@@

We expect to see increased linguistic cohesion among people participating in
online discussions about an idea, event, individual, or movement. While the overall frequency of social media messages related to the event may increase as a result of the event, their linguistic cohesion helps us understand if denizens are using the platforms in a coordinated way that amplifies their messages or goals and creates a virtual protest space. Given that
Syrian protesters coordinated their activity through Twitter by discussing their
own events (e.g., planned protests, active demonstrations) and their reactions
to events happening around them (e.g., government reactions, international
action or inaction), the cohesion in social media should correlate with the
social cohesion of the group [@zeitzoff2011using].

@@

## The Present Work
Social media has become a primary platform for social movements and activism
across the world, but the extent to which virtual activism reflects and drives
real-world mobilization is still largely unknown. In the current work, we
conceptualize the physical and virtual movements as a dynamical system, both
constantly interacting and influencing the other. We aim to study this by
analyzing the real-world events and online social cohesion during nearly three
months of the Arab Spring. Both real- and virtual-world collective coordination
were powerful tools in the events of the Arab Spring and are undeniably linked.
This study aims to explore the dynamics of Twitter users in Syria, who used the
social media platform to raise awareness about the Arab Spring and the relationship of those messages to mass mobilization in the real world.

@@

This work is a reproducible manuscript. Code and analyses can be found at:
www.github.com/mchiovaro/arab-spring.
